---
output: 
  md_document:
  variant: markdown_github
title: ""
author: ""
date: ""
---

# Translating literature-reported quantitative pharmacology models for simulation and interactive visualization

- 20 July 2017 Cambridge Innovation Center, Cambridge MA
- Presentation slide deck: [index.pdf](index.pdf)

# References
The goal of this presentation was to translate and use models reported in the literature. Accordingly, we note that we were not involved in 
developing any of the models presented here, but acknowledge the authors who
not only did the scientific work, but published that work in a way that 
allowed us to both read about and use their models.

<a name="epo"></a>
__Population Pharmacokinetic and Pharmacodynamic Model- Based Comparability Assessment of a Recombinant Human Epoetin Alfa and the Biosimilar HX575__

- Xiaoyu Yan, MS, Philip J. Lowe, PhD, Martin Fink, PhD, Alexander Berghout, PhD, Sigrid Balser, PhD, and Wojciech Krzyzanski, PhD

- J Clin Pharmacol. 2012 November ; 52(11): 1624–1644. doi:10.1177/0091270011421911

- [Model](content/model/epo.cpp), [Example](#epoex), [R script](content/epo.R)

<a name="ddi"></a>
__Quantitative Analyses of Hepatic OATP-Mediated Interactions Between Statins and Inhibitors Using PBPK Modeling With a Parameter Optimization Method__

- T Yoshikado, K Yoshida, N Kotani, T Nakada, R Asaumi, K Toshimoto, K Maeda, H Kusuhara and Y Sugiyama

- CLINICAL PHARMACOLOGY & THERAPEUTICS | VOLUME 100 NUMBER 5 | NOVEMBER 2016 doi:10.1002/cpt.391

- [Model](content/model/yoshikado.cpp), [Example](#ddiex), [R simulation script](content/yoshikado.R), 
[Fit with minqa & RcppDE](content/fit.R), [Fit with MCMCpack](content/fit_mcmc.R)


<a name="erki"></a>
__Clinical responses to ERK inhibition in BRAF{V600E}-mutant colorectal cancer predicted using a computational model__

- Daniel C. Kirouac, Gabriele Schaefer, Jocelyn Chan, Mark Merchant, Christine Orr, Shih-Min A. Huang, John Moffat, Lichuan Liu, Kapil Gadkar and Saroja Ramanujan

- npj Systems Biology and Applications  (2017) 3:14  ; doi:10.1038/s41540-017-0016-1

- [Model](content/model/mapk.cpp), [Example](#erkiex), [R script](content/mapk.R), [Generate figure 6b](content/mapk_figure.R)


# Examples

```{r,echo=FALSE,message=FALSE}
library(dplyr)
library(mrgsolve)
library(ggplot2)
options(mrgsolve_mread_quiet=TRUE)
knitr::opts_chunk$set(comment='.',fig.path="content/img/README-",
                      fig.align="center", warning=FALSE, fig.height=4.5, fig.width=10)
```

<a name="epoex"></a>

## EPO PK/PD model ([Yan et al.](#epo)) [[code](content/model/epo.cpp)]
```{r}
mod <- mread("epo", "content/model")
```

### QW dosing is equally effective as TIW dosing for SC but not IV administration

SC dosing

```{r}
tiw <- ev_days(ev(amt=7800,ID=1,rate=-2), days="m,w,f", addl=3)
qd <- filter(tiw, time==0) %>% mutate(amt=40000,ID=2)

data_sc <- bind_rows(tiw,qd)
```

```{r}
mod %>% zero_re %>% mrgsim(data=data_sc, end=700, delta=0.5) %>% 
  plot(EPOi+RBCi~time)
```

IV dosing

```{r}
data_iv <- mutate(data_sc, rate=0, cmt=2)
```


```{r}
mod %>% zero_re %>% mrgsim(data=data_iv, end=700, delta=0.5) %>% 
  plot(EPOi+RBCi~time)
```

<a name="ddiex"></a>

## OATP-DDI model ([Yoshikado et al.](#ddi)) [[code](content/model/yoshikado.cpp)]

```{r}
mod <- mread("yoshikado", "content/model")
```

Simulate the DDI
```{r}
ddi <- c(ev(amt=2000, cmt=2),
         ev(amt=30,   cmt=1))

ddi <- as.data.frame(ddi) %>% mutate(ID=2)
no_ddi <- filter(ddi, amt==30) %>% mutate(ID=1)

data <- bind_rows(no_ddi,ddi)

out <- mod %>% mrgsim(data=data,end=14, delta=0.1)
```

Plot (code not shown)

```{r,echo=FALSE, fig.width=7}
out <- as.data.frame(out)
ggplot(out, aes(x=time,y=CP,col=factor(ID))) + 
  scale_y_continuous(trans="log10", limit=c(0.1,100), breaks=c(0.1,1,10,100),
                     name="Pitavastatin concentration (ng/mL)") + 
  geom_line(lwd=1) + 
  scale_x_continuous(breaks=seq(0,14,2),limit=c(0,14), name="Time (hours)") + 
  scale_color_manual(values=c("darkslateblue", "firebrick"), 
                     name="", labels=c("No-DDI", "DDI")) + 
  theme(legend.position="top")

```

<a name="erkiex"></a>

## ERK inhibition in CRC ([Kirouac et al.](#erki)) [[code](content/model/mapk.cpp)]

```{r}
mod <- mread("mapk", "content/model")
vp <- read.csv("content/data/s10vpop.csv", header=TRUE) %>% slice(1)
mod <- param(mod,vp) %>% init(vp)
```


Simulate to cycles of  `GDC-0994` dosing

```{r}
data <- expand.ev(amt=400, cmt=12, time=c(0,28), ii=1, addl=20) %>% mutate(ID=1)

out <- mrgsim(mod, data=data, end=56, delta=0.1, Req="GDC,TUMOR")
```

Plot (code not shown)

```{r,echo=FALSE}
sims <- tidyr::gather(as.data.frame(out), variable, value, c(GDC:TUMOR))
sims <- mutate(sims, variable=factor(variable,
                                     labels=c("GDC-0994 PK", "Tumor")))

ggplot(data=sims) + 
  geom_line(aes(time,value,col=variable), lwd=1) + 
  scale_color_manual(values=c("darkslateblue", "firebrick"),guide=FALSE) +
  scale_x_continuous(name="Time (days)", breaks=seq(0,56,14)) + 
  facet_wrap(~variable, scales="free_y")
```


# Implementation details

## R packages
- `dplyr`
- `readr`
- `MCMCpack`
- `minqa`
- `tidyr`
- `ggplot2`
- `purrr`
- `purrlyr`
- `mrgsolvetk`
- `dmutate`


## Special installation instructions
Please be sure to install `mrgsolve`, `dmutate` and `mrgsolvetk` from GitHub:
```r
devtools::install_github("metrumresearchgroup/mrgsolve")
devtools::install_github("mrgsolve/mrgsolvetk")
devtools::install_github("kylebmetrum/dmutate")
```
