---
output: 
  md_document:
    variant: markdown_github
title: ""
author: ""
date: ""
---

# Presentation materials for July 2017 CIC event

- 20 July 2017 Cambridge Innovation Center, Cambridge MA
- Presentation slide deck: [index.pdf](index.pdf)

# References
The goal of this presentation was to translate and use models reported in the literature. Accordingly, we note that we were not involved in 
developing any of the models presented here, but acknowledge the authors who
not only did the scientific work, but published that work in a way that 
allowed us to not only read about their models but to also use those models.

__Population Pharmacokinetic and Pharmacodynamic Model- Based Comparability Assessment of a Recombinant Human Epoetin Alfa and the Biosimilar HX575__

- Xiaoyu Yan, MS, Philip J. Lowe, PhD, Martin Fink, PhD, Alexander Berghout, PhD, Sigrid Balser, PhD, and Wojciech Krzyzanski, PhD

- J Clin Pharmacol. 2012 November ; 52(11): 1624–1644. doi:10.1177/0091270011421911.

- [Model](content/model/epo.cpp), [R script](content/epo.R)

__Quantitative Analyses of Hepatic OATP-Mediated Interactions Between Statins and Inhibitors Using PBPK Modeling With a Parameter Optimization Method__

- T Yoshikado, K Yoshida, N Kotani, T Nakada, R Asaumi, K Toshimoto, K Maeda, H Kusuhara and Y Sugiyama

- CLINICAL PHARMACOLOGY & THERAPEUTICS | VOLUME 100 NUMBER 5 | NOVEMBER 2016 doi:10.1002/cpt.391

- [Model](content/model/yoshikado.cpp), [R script](content/yoshikado.R), 
  [Fit with minqa & RcppDE](content/fit.R), [Fit with MCMCpack](content/fit_mcmc.R)

__Clinical responses to ERK inhibition in BRAF{V600E}-mutant colorectal cancer predicted using a computational model__

- Daniel C. Kirouac, Gabriele Schaefer, Jocelyn Chan, Mark Merchant, Christine Orr, Shih-Min A. Huang, John Moffat, Lichuan Liu, Kapil Gadkar and Saroja Ramanujan

- npj Systems Biology and Applications  (2017) 3:14  ; doi:10.1038/s41540-017-0016-1

- [Model](content/model/mapk.cpp), [R script](content/mapk.R), [Generate figure 6b](content/mapk_figure.R)


# Examples

```{r,echo=FALSE}
library(dplyr)
library(mrgsolve)
options(mrgsolve_mread_quiet=TRUE)
knitr::opts_chunk$set(comment='.',fig.path="content/img/README-")
```


## EPO Model (Yan et al.)
```{r}
mod <- mread("epo", "content/model")
```

### QW dosing is equally effective as TIW dosing for SC but not IV administration

SC dosing
```{r}
tiw <- ev_days(ev(amt=7800,ID=1,rate=-2), days="m,w,f", addl=3)
qd <- filter(tiw, time==0) %>% mutate(amt=40000,ID=2)

data_sc <- bind_rows(tiw,qd)
```

```{r}
mod %>% zero_re %>% mrgsim(data=data_sc, end=700, delta=0.5) %>% plot
```
IV dosing
```{r}
data_iv <- mutate(data_sc, rate=0, cmt=2)
```


```{r}
mod %>% zero_re %>% mrgsim(data=data_iv, end=700, delta=0.5) %>% plot
```

# Implementation details

## R packages
- `dplyr`
- `readr`
- `MCMCpack`
- `minqa`
- `tidyr`
- `ggplot2`

## mrgsolve installation
Please be sure to install `mrgsolve` from GitHub:
```r
devtools::install_github("metrumresearchgroup/mrgsolve")
```
